Master thesis

### 2017-06-09 docs/master-thesis.tex
A2.LL 2

We gathered data from TRY database to evaluate and improve previous
models. We requested functional traits available in Bridge dataset and
leaf lifespan for tropical species present in GLOPNET dataset and used
in TROLL (see \ref{tab:A2traits}). We compiled GLOPNET, TRY, and DRYAD
databases. We kept dataset given by GLOPNET as origin dataset for
observations. It correspond to leaf lifespan and also to LMA and Nmass
in most of cases. We measured variable importance in functionnal traits
to explain leaf lifespan with out-of bag method applied on a random
forest.

@@

We tested different models and we kept the model with the best tradeoff
between model complexity, convergence, likelihood, and prediction
quality (see \href{LL2.html}{leaf lifespan model} document). We found
following model, named \textbf{M1}:
\[M1~~ {LL_s}_j \sim \mathcal{logN}(\beta_0*e^{{\beta_1}_s*{LMA_s}_j - {\beta_2}_s*{Nmass_s}_j},\,\sigma)\,\]
\[s=1,...,S_{=4}~, ~~j=1,...,n_s\]
\[{\beta_i}_s \sim \mathcal{N}({\beta_i},\,\sigma_i)\,^I, ~~(\beta_i, \sigma, \sigma_i) \sim \mathcal{\Gamma}(0.001,\,0.001)\,^{2I+1}\]

Leaf lifespan (\(LL\)) is following a lognormal law with location
\({\mu_s}_j\) and scale \(\sigma\). \(s\) represents the dataset used to
do the fit between 1 and \(S=4\), it encompass environmental and
protocol variations. \(j\) represents the observation for a given site
\(s\). Location \({\mu_s}_j\) is the exponential from the difference
between LMA and Nmass with parameters \({\beta_0}_s\), \({\beta_1}_s\),
and \({\beta_2}_s\). Each \({\beta_i}_s\) is following a normal law
located on \(\beta_i\) with scale \(\sigma_i\). Finally all \(beta_i\),
\(\sigma\), and \(\sigma_i\) are assumed without presemption following
gamma laws of parameters \(10^-2\), \(10^-2\).

We decided to add to model \textbf{M1}, wood specific gravity as a new
predictor through a parameter \(\beta_3\) in a new model named
\textbf{M10}. We removed dataset random effect on \(\beta_3\) due to the
fact that all wood specific gravities were provided by DRYAD database.
Due to a correlation between \(\beta_0\) and \(\beta_3\) in \textbf{M10}
(see figure 5), we tried also a model \textbf{M11} without intercept
\(\beta_0\). We then sampled all models with this composite dataset, and
obtained following results :

@@

\begin{longtable}[]{@{}lrr@{}}
\caption{\label{tab:OOB}Variable importance calculated with out-of the bag
method applied on a random forest. Leaf lifespan (LL) is taken in
GLOPNET database from \citet{wright_worldwide_2004}. Leaf mass per area
(LMA), and leaf nitrogen content (Nmass) are taken both in TRY
(\url{https://www.try-db.org}) and GLOPNET databases. Wood specific
gravity (wsg) is taken both in TRY and DRYAD databases.}\tabularnewline
\toprule
& \%IncMSE & IncNodePurity\tabularnewline
\midrule
\endfirsthead
\toprule
& \%IncMSE & IncNodePurity\tabularnewline
\midrule
\endhead
LMA & 99.69390 & 2028.079\tabularnewline
Nm & 159.03360 & 2666.670\tabularnewline
wsg & 50.97284 & 1475.023\tabularnewline
\bottomrule
\end{longtable}

Table 3: Model fit results for composite dataset. \emph{Data are taken
from GLOPNET, TRY, and DRYAD datasets (see previous description). ML
stand for maximum likelihood and RMSEP for root mean square error of
prediction}

All model have correctly converged.

Model \textbf{M10} have shown a correlation between parameters
\(\beta_0\) and \(\beta_3\), responsible for the creation of model
\textbf{M11} (see previous paragraphs). Model \textbf{M11} only show a
slight link between \(\beta_1\) and \(\beta_3\) which can be considered
as non significant.

We can denote that retained vector of parameters with maximum likelihood
for model \textbf{M11} is close to 95\% confidence limit.

We finally tested the new allometry obtained with the new model
\textbf{M} on TROLL data (figure 7) and with TROLL simulations. Minimum
leaf lifespan is now above 3 months for the new model \textbf{M} with
TROLL species dataset. Moreover lowest leaf lifespan species are not the
one with the smallest LMA with new model. Lastly, new model \textbf{M}
variation of leaf lifespan is almost spanning all previous allometries
prediciton on LMA axis.

Simulations are validating that this new allometry resolve the issue of
unrealistically low aboveground biomass for low LMA species due to an
early edath of individuals inside simulations. For instance with this
allometry Symphonia sp 1 (a low LMA species) is now reaching a realistic
aboveground biomass above X tonnes of C/ha and realistic diameter and
age distribution inside the final population.

>>>>

We compiled functional traits from GLOPNET
\citep{wright_worldwide_2004}, TRY \citep{Kattge2011}, and DRYAD
\citep{chave_towards_2009} databases (see \ref{tab:A2traits}). We kept
dataset given by GLOPNET as origin dataset for observations. Dataset
defined as origin corresponded to leaf lifespan (\(LL\)) and most of the
time to leaf mass per area (\(LMA\)) and leaf nitrogen content per leaf
dry mass (\(Nmass\)). We measured variable importance in functionnal
traits to explain leaf lifespan with out-of bag method applied on a
random forest. Then, we used a béayesian apporach to test different
models with growing level of complexity. We retained the model with the
best tradeoff between model complexity (number of parameters \(K\)),
convergence, likelihood, and prediction quality (root mean sqaure error
of prediciton \(RMSEP\)). We finally tested the new allometry obtained
with the selected model with TROLL simulations.

@@

@@

Out of the bag method applied on a random forest highlighted the
importance of leaf nitrogen content per leaf dry mass (\(N_m\)) to model
leaf lifespan (see \ref{tab:A2OOB}). \(N_m\) importance was higher than
leaf mass per area (158 against 96 percent of mean square error
increase) which was used as a proxy for leaf lifespan in previous
models. Finally, wood specific gravity (\(wsg\)) add also a significant
importance in leaf lifespan estimation.

\begin{longtable}[]{@{}lrr@{}}
\caption{\label{tab:A2OOB}Variable importance calculated with out-of the bag
method applied on a random forest. First column represents the mean
decrease in mean square error (\%IncMSE) whereas second column
represents the total decrease in node impurities, measured by the Gini
Index (IncNodePurety). Leaf lifespan (LL) is taken in GLOPNET database
from \citet{wright_worldwide_2004}. Leaf mass per area (LMA), and leaf
nitrogen content (Nmass) are taken both in TRY
(\url{https://www.try-db.org}) and GLOPNET databases. Wood specific
gravity (wsg) is taken both in TRY and DRYAD databases.}\tabularnewline
\toprule
& \%IncMSE & IncNodePurity\tabularnewline
\midrule
\endfirsthead
\toprule
& \%IncMSE & IncNodePurity\tabularnewline
\midrule
\endhead
LMA & 99.69390 & 2028.079\tabularnewline
Nm & 159.03360 & 2666.670\tabularnewline
wsg & 50.97284 & 1475.023\tabularnewline
\bottomrule
\end{longtable}

The selected model had a maximum likelihood of 13.6 and a RMSEP of 12
months:

Leaf lifespan \(LL\) follows a lognormal law with location infered from
leaf lifespan \(LMA\), nitrogen content \(N\) and wood specific gravity
\(wsg\) and a scale \(\sigma\). Each \({\beta_i}_d\) is following a
normal law located on \(\beta_i\) with a scale of \(\sigma_i\). All
\(\beta_i\), \(\sigma_i\), and \(\sigma\) are assumed without
presemption following a gamma law. \(d\) represents the dataset random
effects and encompass environmental and protocol variations.

\begin{figure}[htbp]
\centering
\includegraphics{master-thesis_files/figure-latex/A2LLpred-1.pdf}
\caption{\label{fig:A2LLpred}Leaflifespan predictions for the selected model
with leaf mass per area (LMA), leaf nitrogen content (Nmass), wood
specific gravity (wsg) and predicted versus observed values. Leaf
lifespan (LL) is predicted with model M10 fit. Leaf mass per area (LMA)
and leaf nitrogen content (Nmass), and wood specific gravity (wsg) are
taken in a composite dataset of GLOPNET, TRY and DRYAD datasets.Warning
LMA (resp. Nmass and wsg) is not constant and depend on the closest
point value for right (resp. center and left) graph.}
\end{figure}

Simulations are validating that this new allometry resolve the issue of
unrealistically low aboveground biomass for low LMA species due to an
early edath of individuals inside simulations. For instance with this
allometry Symphonia sp 1 (a low LMA species) is now reaching a realistic
aboveground biomass above 400 \(tonC.ha^{-1}\) and realistic diameter
and age distribution inside the final population.
